### 2015-07-09 text/ms.Rmd
More ms edits

\noindent
Sean C. Anderson^1\*^, 
Jamie Afflerbach^2^, 
Mark Dickey-Collas^3^, 
Olaf P. Jensen^4^, 
Kristin M. Kleisner^5^, 
Catherine Longo^2^, 
C\'{o}il\'{i}n Minto^6^, 
Giacomo Chato Osio^7^, 
Dan Ovando^8^, 
Andrew A. Rosenberg^9^, 
Elizabeth R. Selig^10^, 
James T. Thorson^11^,
Jessica C. Walsh^1^,
Andrew B. Cooper^1^
*(order to be determined and others may be added)*

@@

Ecological resource managers are often faced with reconciling multiple potentially
conflicting estimates of population status and trajectory.  One solution is to
average these predictions in an ensemble, but this approach ignores covariance
between models and may not optimally reduce bias and improve accuracy.
Superensemble models, commonly used in the fields of climate and weather
science, may provide a superior solution. Superensembles use the predictions
from multiple models as covariates in a new statistical model. Here we evaluate
the potential for ensemble and superensemble models ('ensemble methods') to improve estimates of fish population
status and trajectory. We fit four data-limited models of population and exploitation
dynamics to marine fish stocks and combine their estimates of the mean and
slope of biomass at maximum sustainable yield ($B/B_\mathrm{MSY}$) with an
ensemble average and three superensembles: a linear model, a random forest, and
a boosted regression tree. 
We selected data-limited models for their applicability across the majority of stocks world wide.
We build our models on a simulated dataset of 5760
fish stocks and tested them with cross-validation and a global database of 
`r ram_stocks_n` stock assessments that have also been assessed with more data-intensive stock assessment models.
We find that ensemble methods substantially improve estimates of population status and trajectory.
Machine-learning superensembles performed the best for population status: accuracy improved
`r mean_sim$mach_vs_ind_mare_fold`%, correlation between predicted and true
status improved from 
`r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median
proportional error) declined from 
`r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. 
We found similar improvements when predicting the trajectory of status. When the
simulation-trained superensembles were applied to catch data for global fish stocks, the best
superensembles improved all dimensions of performance over the next-best
individual model. Ensemble methods have the potential to
improve estimates of population status and trajectory across many taxa beyond
fish. However, they must be carefully tested, formed from a diverse set of
accurate models, and built on a dataset sufficiently representative of the
populations to which they are applied.

<!-- up to 12 keywords -->
\noindent 
Keywords: data-limited fisheries, ensemble methods, model averaging, population dynamics, sustainable resource management

>>>>

\noindent
Sean C. Anderson^1\*^, 
Andrew B. Cooper^1^
Olaf P. Jensen^4^, 
C\'{o}il\'{i}n Minto^6^, 
James T. Thorson^11^,
Jessica C. Walsh^1^,
Jamie Afflerbach^2^, 
Mark Dickey-Collas^3^, 
Kristin M. Kleisner^5^, 
Catherine Longo^2^, 
Giacomo Chato Osio^7^, 
Dan Ovando^8^, 
Andrew A. Rosenberg^9^, 
Elizabeth R. Selig^10^, 
*(order to be determined and superscripts out of order)*

@@

Ecological resource managers are often faced with reconciling multiple potentially
conflicting estimates of population status and trajectory.  One solution is to
average these predictions in an ensemble, but this approach ignores covariance
between models and may not optimally reduce bias and improve accuracy.
Superensemble models, commonly used for climate and weather
forecasting, may provide a superior solution. Superensembles use the predictions
from multiple models as covariates in a new statistical model. Here we evaluate
the potential for ensemble and superensemble models ('ensemble methods') to improve estimates of fish population
status and trajectory. We fit four data-limited models of population and exploitation
dynamics to marine fish stocks and combine their estimates of the mean and
slope of biomass at maximum sustainable yield ($B/B_\mathrm{MSY}$) with an
ensemble average and three superensembles: a linear model, a random forest, and
a boosted regression tree. 
We selected data-limited models for their applicability across the majority of stocks world wide.
We build our models on a simulated dataset of 5760
fish stocks and tested them with cross-validation and a global database of 
`r ram_stocks_n` stock assessments that have also been assessed with more data-intensive stock assessment models.
We find that ensemble methods substantially improve estimates of population status and trajectory.
Machine-learning superensembles performed the best for population status: accuracy improved
`r mean_sim$mach_vs_ind_mare_fold`%, correlation between predicted and true
status improved from 
`r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median
proportional error) declined from 
`r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. 
We found similar improvements when predicting the trajectory of status. When the
simulation-trained superensembles were applied to catch data for global fish stocks, the best
superensembles improved all dimensions of performance over the next-best
individual model. Ensemble methods have the potential to
improve estimates of population status and trajectory across many taxa beyond
fish. However, they must be carefully tested, formed from a diverse set of
accurate models, and built on a dataset sufficiently representative of the
populations to which they are applied.

\bigskip
<!-- up to 12 keywords -->
\noindent 
Keywords: data-limited fisheries, ensemble methods, multi-model averaging, population dynamics, sustainable resource management
